---
title: "Conjugate Sampler - Testing Posterior Summaries"
author: "Jonathan Klus"
date: "2024-01-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load necessary functions
source("./Multivariate_DPMM_unknownvar_DEV.R")
source("./Multivariate_DPMM_unknownvar_DEE.R")
source("./Multivariate_DPMM_unknownvar_UVV.R")
source("./posterior_helper_fxns.R")
source("./post_processing_inf.R")

# load R libraries
library(ggplot2)
library(gridExtra)
library(plotly)
library(label.switching)
library(LaplacesDemon)
library(parallel)
library(stringr)
```

## DEE Sampler Setup

- $y_1, \dots, y_n \sim \sum_{k=1}^K\pi_kN(\mu_k, \sigma^2 I_J)$
- $\mu_1, \dots, \mu_K \sim N(\mu_0, r\sigma^2 I_J)$
- $\sigma^2 \sim IG(a,b)$
- $r \sim IG(g, h)$
- $b \sim Ga(d, f)$

## DEV Sampler Setup

- $y_1, \dots, y_n \sim \sum_{k=1}^K\pi_kN(\mu_k, \sigma^2_k I_J)$
- $\mu_1, \dots, \mu_K \sim N(\mu_0, r\sigma^2_k I_J)$
- $\sigma^2_1,\dots, \sigma^2_K \sim IG(a,b)$
- $r \sim IG(g, h)$
- $b \sim Ga(d, f)$

## Sampler Examples

### 2D Example - well separated

Mixture of 3 multivariate normal densities all with diagonal variance structure
and $\sigma^2 = 10$. The true means are $(-20,20)$, $(20,-20)$, and $(0,0)$, and
$\sum_{j}n_j = 30$.

```{r, echo=FALSE}
################## simulate data ##############################################

set.seed(516)
nreps = 10
seeds = sample(x = 1:10^4, size = nreps, replace = FALSE)

w = c(0.4, 0.3, 0.3)
means = list(
  c(-20, 20),
  c(20, -20),
  c(0, 0)
)

var = diag(10, length(means[[1]])) # variances known, diagonal, and equal

yreps = lapply(X = 1:nreps, 
           FUN = function(x){
             set.seed(seeds[x])
             assign = sample(x = 1:length(means), size = 30, replace = TRUE, prob = w)
             y = lapply(X = assign,
                        FUN = function(x){
                        t(mvtnorm::rmvnorm(n = 1, mean = means[[x]], sigma = var))
                          })
             return(list(y = y, assign = assign))
           })

# make a grid of plots
pltlist = list()
for(rep in 1:nreps){

  y_matrix = matrix(data = unlist(yreps[[rep]]$y), ncol = 2, byrow = TRUE)
  assign = yreps[[rep]]$assign

  data = data.frame(
    assign = assign,
    y1 = y_matrix[,1],
    y2 = y_matrix[,2]
  )

  p = ggplot(data = data, aes(x = y1, y = y2)) + #, label = rownames(data))) +
    geom_point(color = assign, size = 0.75) +
    #geom_text(size = 3, hjust = 0, nudge_x = 0.5, color = assign) +
    # geom_text(size = 3, color = assign) +
    theme_classic()

  pltlist[[rep]] = p
}

do.call(grid.arrange, pltlist)#, top = "Simulated 2D Data and True Group Assignments")

# drop params used to generate data, except for var which is assumed known
mu_true = means
var_true = var
# yreps # 2nd element is assign
rm(w, means)
```

#### DEV sampler 

##### without split-merge step

```{r}
# n_iter = 5000
# use_cores = min(parallel::detectCores(), nreps)
# sim_results_DEV_no_sm = parallel::mclapply(X = 1:10, mc.cores = use_cores,
#       FUN = function(x){MVN_CRP_sampler_DEV(
#         S = n_iter, seed = seeds[x], y = yreps[[x]]$y,
#         alpha = 1, r = 10, g = 1, h = 50,
#         sigma_hyperprior = FALSE, fix_r = FALSE,
#         mu0 = matrix(round((colMeans(matrix(unlist(yreps[[x]]$y), ncol = 2))),0), ncol = 1),
#         a = 1, b = 50,
#         k_init = 1, diag_weights = FALSE,
#         verbose = FALSE, split_merge = FALSE
#         )}
#       )
# 
# 
# datafile_name = paste0("../MCMC_Runs/conjDEVsamp_minisimstudy_noSM_", 
#                        stringr::str_replace_all(Sys.Date(), pattern = "-", replacement = "_"), ".rds")
# saveRDS(object = sim_results_DEV_no_sm, file = datafile_name)
sim_results_DEV_no_sm = readRDS(file = "../MCMC_Runs/conjDEVsamp_minisimstudy_noSM_2024_01_16.rds")

DEV_no_sm_sum = dpmm_summary(output = sim_results_DEV_no_sm, 
                             dataset_ind = 1, 
                             print_result = TRUE, 
                             make_traceplot = TRUE,
                             burn_in = 1000, t_hold = 100, 
                             num_dims = 2)

# for(rep in 1:nreps){
#   print(table(sim_results_DEV_no_sm[[rep]]$k))
#   p1 = make_k_traceplot(k = sim_results_DEV_no_sm[[rep]]$k, 
#                         group_assign = sim_results_DEV_no_sm[[rep]]$group_assign)
#   print(p1)
# }
```


##### with split-merge step

```{r}
# sim_results_DEV_with_sm = parallel::mclapply(X = 1:10, mc.cores = use_cores,
#       FUN = function(x){MVN_CRP_sampler_DEV(
#         S = n_iter, seed = seeds[x], y = yreps[[x]]$y,
#         alpha = 1, r = 10, g = 1, h = 50,
#         sigma_hyperprior = FALSE, fix_r = FALSE,
#         mu0 = matrix(round((colMeans(matrix(unlist(yreps[[x]]$y), ncol = 2))),0), ncol = 1),
#         a = 1, b = 50,
#         k_init = 1, diag_weights = FALSE,
#         verbose = FALSE, split_merge = TRUE, sm_iter = 10
#         )}
#       )
# 
# datafile_name = paste0("../MCMC_Runs/conjDEVsamp_minisimstudy_withSM_", 
#                        stringr::str_replace_all(Sys.Date(), pattern = "-", replacement = "_"), ".rds")
# saveRDS(object = sim_results_DEV_with_sm, file = datafile_name)
# sim_results_DEV_with_sm = readRDS(file = "../MCMC_Runs/conjDEVsamp_minisimstudy_withSM_6_Dec_2023.rds")

sim_results_DEV_with_sm = readRDS(file = "../MCMC_Runs/conjDEVsamp_minisimstudy_withSM_2024_01_16.rds")

# got an error here with data set 1 that I haven't seen before -- need to investigate
# further
DEV_with_sm_sum = dpmm_summary(output = sim_results_DEV_with_sm, 
                             dataset_ind = 2, 
                             print_result = TRUE, 
                             make_traceplot = TRUE,
                             burn_in = 1000, t_hold = 100, 
                             num_dims = 2)
```

# debugging 

# DEV sample debugging 

```{r}
sim_results_DEV_with_sm = readRDS(file = "../MCMC_Runs/conjDEVsamp_minisimstudy_withSM_2024_01_16.rds")

# got an error here with data set 1 that I haven't seen before -- need to investigate
# further
DEV_with_sm_sum = lapply(X = 1:length(sim_results_DEV_with_sm), 
                         FUN = function(x){
                           cat("\n Dataset #", x)
                           dpmm_summary(output = sim_results_DEV_with_sm, 
                             dataset_ind = x, 
                             print_result = TRUE, 
                             make_traceplot = TRUE,
                             burn_in = 1000, t_hold = 100, 
                             num_dims = 2)
                           })

output = readRDS(file = "../MCMC_Runs/conjDEVsamp_minisimstudy_withSM_2024_01_16.rds")
dataset_ind = 8
print_result = TRUE
burn_in = 1000
t_hold = 100
num_dims = 2

# table(apply(X = output[[1]]$group_assign, MARGIN = 1, FUN = function(x){length(unique(x))}))

prob_list_by_k = get_probs_by_k(probs = output[[dataset_ind]]$group_probs,
                                n_groups = output[[dataset_ind]]$k,
                                burn_in = burn_in,
                                iter_threshold = t_hold)
group_assign_list_by_k = get_assign_by_k(assign = output[[dataset_ind]]$group_assign,
                                         n_groups = output[[dataset_ind]]$k,
                                         burn_in = burn_in,
                                         iter_threshold = t_hold)

# correct label switching
stephens_result = get_stephens_result(group_assign_list_by_k = group_assign_list_by_k,
                                      prob_list_by_k = prob_list_by_k$prob_list)

# summarize means & variances
mean_list_by_k_stephens = list_params_by_k(draws = output[[dataset_ind]]$means,
                                           # k_vec = output[[1]]$k,
                                           # burn_in = burn_in,
                                           # iter_threshold = thold,
                                           iter_list = prob_list_by_k$iter_list,
                                           relabel = TRUE,
                                           permutation = stephens_result,
                                           param_type = "Mean")

var_list_by_k_stephens = list_params_by_k(draws = output[[dataset_ind]]$vars,
                                          iter_list = prob_list_by_k$iter_list,
                                          relabel = TRUE,
                                          permutation = stephens_result,
                                          param_type = "Var")

calc_KL_diverg(y = output[[1]]$data,
               mu_est = mean_list_by_k_stephens,
               Sigma_est = var_list_by_k_stephens,
               group_assign = stephens_result,
               true_assign = yreps[[1]]$assign,
               mu_true = mu_true,
               Sigma_true = var_true,
               equal_var_assump = TRUE)
```


## UVV sample debugging 

```{r}
sim_results_UVV_no_sm = readRDS(file = "../MCMC_Runs/conjUVVsamp_minisimstudy_noSM_2024_01_16.rds")

# got an error here with data set 1 that I haven't seen before -- need to investigate
# further
UVV_no_sm_sum = dpmm_summary(output = sim_results_UVV_no_sm, 
                            dataset_ind = 10, 
                            print_result = TRUE, 
                            make_traceplot = TRUE,
                            burn_in = 1000, t_hold = 100, 
                            num_dims = 2)
```

