---
title: "KL Divergence Summary"
author: "Jonathan Klus"
date: "2024-02-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(ggplot2)
library(gridExtra)
library(plotly)
library(label.switching)
library(LaplacesDemon)
library(parallel)
library(stringr)
```

$$
D_{KL}(p||q) = \sum_{x \in \mathcal{X}}p(x)log\left(\frac{p(x)}{q(x)}\right)
$$

for densities $p$ and $q$ with support $\mathcal{X}$, where $p$ is the true density and $q$ is the estimated density.

## Well separated case

### Simulated data sets

Mixture of 3 multivariate normal densities all with diagonal variance structure
and $\sigma^2 = 10$. The true means are $(-20,20)$, $(20,-20)$, and $(0,0)$, and
$\sum_{j}n_j = 30$.

```{r, echo=FALSE}
################## simulate data ##############################################

set.seed(516)
nreps = 10
seeds = sample(x = 1:10^4, size = nreps, replace = FALSE)

w = c(0.4, 0.3, 0.3)
means = list(
  c(-20, 20),
  c(20, -20),
  c(0, 0)
)

var = diag(10, length(means[[1]])) # variances known, diagonal, and equal

yreps = lapply(X = 1:nreps, 
           FUN = function(x){
             set.seed(seeds[x])
             assign = sample(x = 1:length(means), size = 30, replace = TRUE, prob = w)
             y = lapply(X = assign,
                        FUN = function(x){
                        t(mvtnorm::rmvnorm(n = 1, mean = means[[x]], sigma = var))
                          })
             return(list(y = y, assign = assign))
           })

# make a grid of plots
pltlist = list()
for(rep in 1:nreps){

  y_matrix = matrix(data = unlist(yreps[[rep]]$y), ncol = 2, byrow = TRUE)
  assign = yreps[[rep]]$assign

  data = data.frame(
    assign = assign,
    y1 = y_matrix[,1],
    y2 = y_matrix[,2]
  )

  p = ggplot(data = data, aes(x = y1, y = y2)) + #, label = rownames(data))) +
    geom_point(color = assign, size = 0.75) +
    #geom_text(size = 3, hjust = 0, nudge_x = 0.5, color = assign) +
    # geom_text(size = 3, color = assign) +
    theme_classic()

  pltlist[[rep]] = p
}

do.call(grid.arrange, pltlist)#, top = "Simulated 2D Data and True Group Assignments")

# drop params used to generate data, except for var which is assumed known
mu_true = means
var_true = var
# yreps # 2nd element is assign
rm(w, means)
```

### Results

```{r}
wellsep_kld = readRDS("../MCMC_Runs/kld_summary_sep_02_08_2024.rds")
knitr::kable(x = wellsep_kld, digits = 3, 
             caption = "KL divergence averaged across 10 simulated data sets for the 
             well-separated case with n=30 observations")
```

## Close together case

### Simulated data sets

Mixture of 3 multivariate normal densities all with diagonal variance structure
and $\sigma^2 = 10$. The true means are $(10,10)$, $(0,-5)$, , $(12,-2)$, and $(0,10)$, and
$\sum_{j}n_j = 50$.

```{r, echo=FALSE}
################## simulate data ##############################################

set.seed(516)
nreps = 10
seeds = sample(x = 1:10^4, size = nreps, replace = FALSE)

w = c(.35, .25, .4)
means = list(
  #c(10, 10),
  c(0, -5),
  c(0, 10),
  c(12,2)
)

var = diag(9, length(means[[1]])) # variances known, diagonal, and equal

yreps = lapply(X = 1:nreps, 
           FUN = function(x){
             set.seed(seeds[x])
             assign = sample(x = 1:length(means), size = 30, replace = TRUE, prob = w)
             y = lapply(X = assign,
                        FUN = function(x){
                        t(mvtnorm::rmvnorm(n = 1, mean = means[[x]], sigma = var))
                          })
             return(list(y = y, assign = assign))
           })

# make a grid of plots
pltlist = list()
for(rep in 1:nreps){

  y_matrix = matrix(data = unlist(yreps[[rep]]$y), ncol = 2, byrow = TRUE)
  assign = yreps[[rep]]$assign

  data = data.frame(
    assign = assign,
    y1 = y_matrix[,1],
    y2 = y_matrix[,2]
  )

  p = ggplot(data = data, aes(x = y1, y = y2)) + #, label = rownames(data))) +
    geom_point(color = assign, size = 0.75) +
    #geom_text(size = 3, hjust = 0, nudge_x = 0.5, color = assign) +
    # geom_text(size = 3, color = assign) +
    theme_classic()

  pltlist[[rep]] = p
}

do.call(grid.arrange, pltlist)#, top = "Simulated 2D Data and True Group Assignments")

# drop params used to generate data, except for var which is assumed known
mu_true = means
var_true = var
# yreps # 2nd element is assign
rm(w, means)
```
\newpage

### Results

```{r}
close_kld = readRDS("../MCMC_Runs/kld_summary_close_02_12_2024.rds")
knitr::kable(x = close_kld, digits = 3, 
             caption = "KL divergence averaged across 10 simulated data sets for the 
             closer case with n=30 observations")
```