---
title: "Non-Conjugate Sampler - No SM Testing"
author: "Jonathan Klus"
date: "2024-04-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load necessary functions
# source("./Multivariate_DPMM_nonconj_DEV.R")
source("./Multivariate_DPMM_nonconj_UVV.R")
source("./posterior_helper_fxns.R")
source("./post_processing_inf.R")

# load R libraries
library(ggplot2)
library(label.switching)
library(LaplacesDemon)
library(mvtnorm)
library(stringr)
```

## Sampler Examples

### 2D Example - well separated

Mixture of 3 multivariate normal densities all with diagonal variance structure
and $\sigma^2 = 10$. The true means are $(-20,20)$, $(20,-20)$, and $(0,0)$, and
$\sum_{j}n_j = 30$.

```{r}
################## simulate data ##############################################

set.seed(516)
nreps = 10
seeds = sample(x = 1:10^4, size = nreps, replace = FALSE)

w = c(0.4, 0.3, 0.3)
means = list(
  c(-20, 20),
  c(20, -20),
  c(0, 0)
)

var = diag(10, length(means[[1]])) # variances known, diagonal, and equal

yreps = lapply(X = 1:nreps, 
           FUN = function(x){
             set.seed(seeds[x])
             assign = sample(x = 1:length(means), size = 30, replace = TRUE, prob = w)
             y = lapply(X = assign,
                        FUN = function(x){
                        t(mvtnorm::rmvnorm(n = 1, mean = means[[x]], sigma = var))
                          })
             return(list(y = y, assign = assign))
           })

# make a grid of plots
# pltlist = list()
# for(rep in 1:nreps){
# 
#   y_matrix = matrix(data = unlist(yreps[[rep]]$y), ncol = 2, byrow = TRUE)
#   assign = yreps[[rep]]$assign
# 
#   data = data.frame(
#     assign = assign,
#     y1 = y_matrix[,1],
#     y2 = y_matrix[,2]
#   )
# 
#   p = ggplot(data = data, aes(x = y1, y = y2)) + #, label = rownames(data))) +
#     geom_point(color = assign, size = 0.75) +
#     #geom_text(size = 3, hjust = 0, nudge_x = 0.5, color = assign) +
#     # geom_text(size = 3, color = assign) +
#     theme_classic()
# 
#   pltlist[[rep]] = p
# }
# 
# do.call(grid.arrange, pltlist)#, top = "Simulated 2D Data and True Group Assignments")
# 
# # drop params used to generate data, except for var which is assumed known
# rm(w, means)
```

#### DEV sampler 

##### without split-merge step

```{r}
n_iter = 12000
x = 5 
y_matrix = matrix(data = unlist(yreps[[x]]$y), ncol = 2, byrow = TRUE)
mod_test1 = MVN_CRP_nonconj_DEV(
        S = n_iter, seed = seeds[x], y = yreps[[x]]$y, alpha = 1, 
        mu0 = matrix(round((colMeans(matrix(unlist(yreps[[x]]$y), ncol = 2))),0), ncol = 1),
        sigma0 = 100,
        a = 1, b = 10, sigma_hyperprior = FALSE,
        k_init = 1, diag_weights = FALSE,
        verbose = TRUE, print_iter = 1000, split_merge = FALSE
        )

table(mod_test1$k)
```

```{r, eval=FALSE, include = FALSE}
n_iter = 12000
x = 5 # first simulated data set 
y_matrix = matrix(data = unlist(yreps[[x]]$y), ncol = 2, byrow = TRUE)
mod_test2 = MVN_CRP_nonconj_UVV(
        S = n_iter, seed = seeds[x], y = yreps[[x]]$y, alpha = 1,
        mu0 = matrix(round((colMeans(matrix(unlist(yreps[[x]]$y), ncol = 2))),0), ncol = 1),
        Sigma0 = diag(50, 2), Lambda0 = diag(10,2), nu = 5,
        k_init = 1, diag_weights = FALSE, 
        truth = list(mu_true = means, var_true = var, assign_true = yreps[[x]]$assign), 
        verbose = TRUE, print_iter = 100, split_merge = FALSE
        )

table(mod_test1$k)

# do summary
var_true = list(var, var, var)
modsum1 = dpmm_summary(output = mod_test1, 
                        # dataset_ind = x, 
                        print_k_sum = TRUE, 
                        # print_phi_sum = TRUE,
                        make_traceplot = FALSE,
                        burn_in = 2000, t_hold = 100, 
                        num_dims = 2, equal_var = FALSE,
                        calc_perf = TRUE, mu_true = means, 
                        var_true = var_true, assign_true = yreps[[x]]$assign)
```

## 2D Example - close together

Mixture of 3 multivariate normal densities all with diagonal variance structure
and $\sigma^2 = 10$. The true means are $(10,10)$, $(0,-5)$, , $(12,-2)$, and $(0,10)$, and
$\sum_{j}n_j = 50$.

```{r, echo=FALSE}
################## simulate data ##############################################
w = c(.35, .25, .4)
means = list(
  #c(10, 10),
  c(0, -5),
  c(0, 10),
  c(12,2)
)

var = diag(9, length(means[[1]])) # variances known, diagonal, and equal

yreps = lapply(X = 1:nreps, 
           FUN = function(x){
             set.seed(seeds[x])
             assign = sample(x = 1:length(means), size = 30, replace = TRUE, prob = w)
             y = lapply(X = assign,
                        FUN = function(x){
                        t(mvtnorm::rmvnorm(n = 1, mean = means[[x]], sigma = var))
                          })
             return(list(y = y, assign = assign))
           })

# # make a grid of plots
# pltlist = list()
# for(rep in 1:nreps){
# 
#   y_matrix = matrix(data = unlist(yreps[[rep]]$y), ncol = 2, byrow = TRUE)
#   assign = yreps[[rep]]$assign
# 
#   data = data.frame(
#     assign = assign,
#     y1 = y_matrix[,1],
#     y2 = y_matrix[,2]
#   )
# 
#   p = ggplot(data = data, aes(x = y1, y = y2)) + #, label = rownames(data))) +
#     geom_point(color = assign, size = 0.75) +
#     #geom_text(size = 3, hjust = 0, nudge_x = 0.5, color = assign) +
#     # geom_text(size = 3, color = assign) +
#     theme_classic()
# 
#   pltlist[[rep]] = p
# }
# 
# do.call(grid.arrange, pltlist)#, top = "Simulated 2D Data and True Group Assignments")
# 
# # drop params used to generate data, except for var which is assumed known
# rm(w, means)
```

#### DEV sampler 

##### without split-merge step

```{r, eval=FALSE}
x = 5 # first simulated data set 
n_iter = 12000
y_matrix = matrix(data = unlist(yreps[[x]]$y), ncol = 2, byrow = TRUE)
mod_test2 = MVN_CRP_nonconj_DEV(
        S = n_iter, seed = seeds[x], y = yreps[[x]]$y, alpha = 1, 
        mu0 = matrix(round((colMeans(matrix(unlist(yreps[[x]]$y), ncol = 2))),0), ncol = 1),
        sigma0 = 25,
        a = 1, b = 10, sigma_hyperprior = FALSE,
        k_init = 1, diag_weights = FALSE,
        verbose = TRUE, print_iter = 1000, split_merge = FALSE
        )


table(mod_test2$k)
```

