---
title: "Non-Conjugate Sampler - No SM Testing"
author: "Jonathan Klus"
date: "2024-04-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load necessary functions
# source("./Multivariate_DPMM_nonconj_DEV.R")
source("./Multivariate_DPMM_nonconj_UVV.R")
source("./posterior_helper_fxns.R")

# load R libraries
library(ggplot2)
library(gridExtra)
library(label.switching)
library(LaplacesDemon)
library(stringr)
```

## Sampler Examples

### 2D Example - well separated

Mixture of 3 multivariate normal densities all with diagonal variance structure
and $\sigma^2 = 10$. The true means are $(-20,20)$, $(20,-20)$, and $(0,0)$, and
$\sum_{j}n_j = 30$.

```{r}
################## simulate data ##############################################

set.seed(516)
nreps = 10
seeds = sample(x = 1:10^4, size = nreps, replace = FALSE)

w = c(0.4, 0.3, 0.3)
means = list(
  c(-20, 20),
  c(20, -20),
  c(0, 0)
)

var = diag(10, length(means[[1]])) # variances known, diagonal, and equal

yreps = lapply(X = 1:nreps, 
           FUN = function(x){
             set.seed(seeds[x])
             assign = sample(x = 1:length(means), size = 30, replace = TRUE, prob = w)
             y = lapply(X = assign,
                        FUN = function(x){
                        t(mvtnorm::rmvnorm(n = 1, mean = means[[x]], sigma = var))
                          })
             return(list(y = y, assign = assign))
           })

# make a grid of plots
pltlist = list()
for(rep in 1:nreps){

  y_matrix = matrix(data = unlist(yreps[[rep]]$y), ncol = 2, byrow = TRUE)
  assign = yreps[[rep]]$assign

  data = data.frame(
    assign = assign,
    y1 = y_matrix[,1],
    y2 = y_matrix[,2]
  )

  p = ggplot(data = data, aes(x = y1, y = y2)) + #, label = rownames(data))) +
    geom_point(color = assign, size = 0.75) +
    #geom_text(size = 3, hjust = 0, nudge_x = 0.5, color = assign) +
    # geom_text(size = 3, color = assign) +
    theme_classic()

  pltlist[[rep]] = p
}

do.call(grid.arrange, pltlist)#, top = "Simulated 2D Data and True Group Assignments")

# drop params used to generate data, except for var which is assumed known
rm(w, means)
```

#### DEV sampler 

##### without split-merge step

```{r}
n_iter = 12000
# use_cores = min(parallel::detectCores(), nreps)
# sim_results_DEV_no_sm = parallel::mclapply(X = 1:10, mc.cores = use_cores,
#       FUN = function(x){MVN_CRP_sampler_DEV(
#         S = n_iter, seed = seeds[x], y = yreps[[x]]$y,
#         alpha = 1, r = 10, g = 1, h = 50,
#         sigma_hyperprior = FALSE, fix_r = FALSE,
#         mu0 = matrix(round((colMeans(matrix(unlist(yreps[[x]]$y), ncol = 2))),0), ncol = 1),
#         a = 1, b = 50,
#         k_init = 1, diag_weights = FALSE,
#         verbose = FALSE, split_merge = FALSE
#         )}
#       )

x = 2 # first simulated data set 
y_matrix = matrix(data = unlist(yreps[[x]]$y), ncol = 2, byrow = TRUE)
mod_test1 = MVN_CRP_nonconj_DEV(
        S = n_iter, seed = seeds[x], y = yreps[[x]]$y, alpha = 1, 
        mu0 = matrix(round((colMeans(matrix(unlist(yreps[[x]]$y), ncol = 2))),0), ncol = 1),
        sigma0 = 50,
        a = 1, b = 10, sigma_hyperprior = FALSE,
        k_init = 1, diag_weights = FALSE,
        verbose = TRUE, print_iter = 1000, split_merge = FALSE
        )

table(mod_test1$k)

saveRDS(object = mod_test1, file = "../MCMC_Runs/nonconjDEV_ds2_test_wellsep.rds")


# datafile_name = paste0("../MCMC_Runs/conjDEVsamp_minisimstudy_noSM_", 
#                        stringr::str_replace_all(Sys.Date(), pattern = "-", replacement = "_"), ".rds")
# saveRDS(object = sim_results_DEV_no_sm, file = datafile_name)
# # sim_results_DEV_no_sm = readRDS(file = datafile_name)
# 
# for(rep in 1:nreps){
#   print(table(sim_results_DEV_no_sm[[rep]]$k))
#   p1 = make_k_traceplot(k = sim_results_DEV_no_sm[[rep]]$k, 
#                         group_assign = sim_results_DEV_no_sm[[rep]]$group_assign)
#   print(p1)
# }
```


## 2D Example - close together

Mixture of 3 multivariate normal densities all with diagonal variance structure
and $\sigma^2 = 10$. The true means are $(10,10)$, $(0,-5)$, , $(12,-2)$, and $(0,10)$, and
$\sum_{j}n_j = 50$.

```{r, echo=FALSE}
################## simulate data ##############################################
w = c(.35, .25, .4)
means = list(
  #c(10, 10),
  c(0, -5),
  c(0, 10),
  c(12,2)
)

var = diag(9, length(means[[1]])) # variances known, diagonal, and equal

yreps = lapply(X = 1:nreps, 
           FUN = function(x){
             set.seed(seeds[x])
             assign = sample(x = 1:length(means), size = 30, replace = TRUE, prob = w)
             y = lapply(X = assign,
                        FUN = function(x){
                        t(mvtnorm::rmvnorm(n = 1, mean = means[[x]], sigma = var))
                          })
             return(list(y = y, assign = assign))
           })

# make a grid of plots
pltlist = list()
for(rep in 1:nreps){

  y_matrix = matrix(data = unlist(yreps[[rep]]$y), ncol = 2, byrow = TRUE)
  assign = yreps[[rep]]$assign

  data = data.frame(
    assign = assign,
    y1 = y_matrix[,1],
    y2 = y_matrix[,2]
  )

  p = ggplot(data = data, aes(x = y1, y = y2)) + #, label = rownames(data))) +
    geom_point(color = assign, size = 0.75) +
    #geom_text(size = 3, hjust = 0, nudge_x = 0.5, color = assign) +
    # geom_text(size = 3, color = assign) +
    theme_classic()

  pltlist[[rep]] = p
}

do.call(grid.arrange, pltlist)#, top = "Simulated 2D Data and True Group Assignments")

# drop params used to generate data, except for var which is assumed known
rm(w, means)
```

#### DEV sampler 

##### without split-merge step

```{r}
x = 2 # first simulated data set 
y_matrix = matrix(data = unlist(yreps[[x]]$y), ncol = 2, byrow = TRUE)
mod_test2 = MVN_CRP_nonconj_DEV(
        S = n_iter, seed = seeds[x], y = yreps[[x]]$y, alpha = 1, 
        mu0 = matrix(round((colMeans(matrix(unlist(yreps[[x]]$y), ncol = 2))),0), ncol = 1),
        sigma0 = 225, # round(max(apply(X = y_matrix, MARGIN = 2, FUN = var)),-1), 
        a = 1, b = 10, sigma_hyperprior = FALSE,
        k_init = 1, diag_weights = FALSE,
        verbose = TRUE, print_iter = 500, split_merge = FALSE
        )

table(mod_test2$k)

# datafile_name = paste0("../MCMC_Runs/conjDEVsamp_minisimstudy_close_noSM_",
#                        stringr::str_replace_all(Sys.Date(), pattern = "-", replacement = "_"), ".rds")
saveRDS(object = mod_test2, file = "../MCMC_Runs/nonconjDEV_ds2_test_close.rds")
# sim_results_DEV_no_sm = readRDS(file = datafile_name)

# for(rep in 1:nreps){
#   print(table(sim_results_DEV_no_sm[[rep]]$k))
#   p1 = make_k_traceplot(k = sim_results_DEV_no_sm[[rep]]$k, 
#                         group_assign = sim_results_DEV_no_sm[[rep]]$group_assign)
#   print(p1)
# }
```

