---
title: "Model Spec Comparison"
author: "Jonathan Klus"
date: "2024-03-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# homemade functions
source("./posterior_helper_fxns.R")

# load R libraries
library(ggplot2)
library(dplyr)
library(stringr)
library(knitr) # kable
library(kableExtra) # use for complex tables http://haozhu233.github.io/kableExtra/awesome_table_in_pdf.pdf 

######################## DEFINE HELPER FUNCTIONS################################
# running locally
# data_path = "//GENE.bst.rochester.edu/Projects/JKSTproj/BlueHive_Sim_Results/"

# running on server
data_path = "/projects/jklus/JKSTproj/UPSTAT"

# dev_withSM = readRDS(paste0(data_path, "/MODSUM_conjDEV_3close_n30_withSM_sim_results_ENAR/sum_1.rds"))

# list file extensions
file_ext = list.dirs(data_path)

# for each extension, make a new element of a data table with simulation attributes
summary_table = data.frame(Model = NA, Scenario = NA, SM = NA, Tag = NA, n_obs = NA, n_datasets = NA, dataset_no = NA,
                           ARI = NA, KL = NA, Time = NA) # should add , SumTime = NA for mod summary compute time

for(ext_index in 2:length(file_ext)){ # skip first one, listing main directory
                                      # skip second one, Archive directory
  # list files
  file_list = list.files(path = file_ext[ext_index])
  if(length(file_list) > 0){
    
    temp_summary_table = data.frame(Model = NA, Scenario = NA, SM = NA, Tag = NA, n_obs = NA, n_datasets = NA, dataset_no = NA,
                                    ARI = NA, KL = NA, Time = NA) # should add , SumTime = NA for mod summary compute time
    # parse out attributes and save
    short_file_ext = stringr::str_split(string = file_ext[ext_index], pattern = "/MODSUM")[[1]][2]
    parsed_dirname = unlist(stringr::str_extract_all(string = file_ext[ext_index], pattern = "_[:alnum:]+"))
    temp_summary_table[1:length(file_list), "Model"] = stringr::str_remove(string = parsed_dirname[1], pattern = "_")
    temp_summary_table[1:length(file_list), "Scenario"] = stringr::str_remove(string = parsed_dirname[2], pattern = "_")
    temp_summary_table[1:length(file_list), "n_obs"] = stringr::str_remove(string = parsed_dirname[3], pattern = "_n")
    temp_summary_table[1:length(file_list), "SM"] = stringr::str_remove(string = parsed_dirname[4], pattern = "_")
    temp_summary_table[1:length(file_list), "Tag"] = stringr::str_remove(string = parsed_dirname[length(parsed_dirname)], pattern = "_")
    temp_summary_table[, "dataset_no"] = as.numeric(stringr::str_remove_all(string = file_list, pattern = "sum_|.rds"))
    
    # loop through all files in directory
    sum_bydataset = t(sapply(X = 1:length(file_list), 
                             FUN = function(x){
                               mod_sum = readRDS(paste0(file_ext[ext_index], "/", file_list[x]))
                               c(mod_sum$mean_ARI, mod_sum$kl_div, as.numeric(mod_sum$fit_runtime))
                             }))
    temp_summary_table[, "ARI"] = sum_bydataset[,1]
    temp_summary_table[, "KL"] = sum_bydataset[,2]
    temp_summary_table[, "Time"] = sum_bydataset[,3]
    summary_table = rbind(summary_table, temp_summary_table)
    
  }
  
}

summary_table = summary_table[-1,] # get rid of first row of NAs 
```

## Standard

- All diagonal models have same prior setup (probably not a good idea)
- $\sigma^2 \sim IG(a=1, b=50)$
- For invWish prior, $\Lambda_0 = diag(15)$

## Computation Time

```{r}

summary_table %>% 
  dplyr::filter(Tag == "ENAR") %>%
  dplyr::group_by(Model, Scenario, SM, n_obs) %>%
  dplyr::summarize(Count = n()) %>%
  tidyr::pivot_wider(names_from = c(n_obs, Scenario), values_from = Count) %>%
  # keep first 2 columns with model info, sort remaining cols with results by n
  dplyr::select(1,2, gtools::mixedorder(names(.)[3:length(names(.))])+2)

# time
summary_table %>% 
  dplyr::filter(Tag == "ENAR") %>%
  dplyr::group_by(Model, Scenario, SM, n_obs) %>%
  dplyr::summarize(Time = round(mean(Time),2)) %>%
  tidyr::pivot_wider(names_from = c(n_obs, Scenario), values_from = Time) %>%
  # keep first 2 columns with model info, sort remaining cols with results by n
  dplyr::select(1,2, gtools::mixedorder(names(.)[3:length(names(.))])+2)
```

### Adj Rand Index

```{r}
# ARI
summary_table %>% 
  dplyr::filter(Tag == "ENAR") %>%
  dplyr::group_by(Model, Scenario, SM, n_obs) %>%
  dplyr::summarize(ARI = round(mean(ARI),2)) %>%
  tidyr::pivot_wider(names_from = c(n_obs, Scenario), values_from = ARI) %>%
  # keep first 2 columns with model info, sort remaining cols with results by n
  dplyr::select(1,2, gtools::mixedorder(names(.)[3:length(names(.))])+2)
```

### KL divergence

```{r}
# KL
summary_table %>% 
  dplyr::filter(Tag == "ENAR") %>%
  dplyr::group_by(Model, Scenario, SM, n_obs) %>%
  dplyr::summarize(KL = round(mean(KL),2)) %>%
  tidyr::pivot_wider(names_from = c(n_obs, Scenario), values_from = KL) %>%
  # keep first 2 columns with model info, sort remaining cols with results by n
  dplyr::select(1,2, gtools::mixedorder(names(.)[3:length(names(.))])+2) 
```


## Alternative 1 for close together

- $\sigma^2 \sim IG(a=1, b=25)$
- For invWish prior, $\Lambda_0 = diag(8)$

### Adj Rand Index

```{r}
# ARI
summary_table %>% 
  dplyr::filter(Tag == "altprior") %>%
  dplyr::group_by(Model, Scenario, SM, n_obs) %>%
  dplyr::summarize(ARI = round(mean(ARI),2)) %>%
  tidyr::pivot_wider(names_from = c(n_obs, Scenario), values_from = ARI) %>%
  # keep first 2 columns with model info, sort remaining cols with results by n
  dplyr::select(1,2, gtools::mixedorder(names(.)[3:length(names(.))])+2)
```

### KL divergence

```{r}
# KL
summary_table %>% 
  dplyr::filter(Tag == "altprior") %>%
  dplyr::group_by(Model, Scenario, SM, n_obs) %>%
  dplyr::summarize(KL = round(mean(KL),2)) %>%
  tidyr::pivot_wider(names_from = c(n_obs, Scenario), values_from = KL) %>%
  # keep first 2 columns with model info, sort remaining cols with results by n
  dplyr::select(1,2, gtools::mixedorder(names(.)[3:length(names(.))])+2) 
```

## Alternative 2 for close together

- $\sigma^2 \sim IG(a=1, b=8)$

### Adj Rand Index

```{r}
# ARI
summary_table %>% 
  dplyr::filter(Tag == "altprior2") %>%
  dplyr::group_by(Model, Scenario, SM, n_obs) %>%
  dplyr::summarize(ARI = round(mean(ARI),2)) %>%
  tidyr::pivot_wider(names_from = c(n_obs, Scenario), values_from = ARI) %>%
  # keep first 2 columns with model info, sort remaining cols with results by n
  dplyr::select(1,2, gtools::mixedorder(names(.)[3:length(names(.))])+2)
```

### KL divergence

```{r}
# KL
summary_table %>% 
  dplyr::filter(Tag == "altprior2") %>%
  dplyr::group_by(Model, Scenario, SM, n_obs) %>%
  dplyr::summarize(KL = round(mean(KL),2)) %>%
  tidyr::pivot_wider(names_from = c(n_obs, Scenario), values_from = KL) %>%
  # keep first 2 columns with model info, sort remaining cols with results by n
  dplyr::select(1,2, gtools::mixedorder(names(.)[3:length(names(.))])+2) 
```

## Alternative 3 for close together - hyperprior

- $\sigma^2 \sim IG(a=1, b)$
- $b \sim Ga(g = 1, h = 5)$

### Adj Rand Index

```{r}
# ARI
summary_table %>% 
  dplyr::filter(Tag == "sigmahyp") %>%
  dplyr::group_by(Model, Scenario, SM, n_obs) %>%
  dplyr::summarize(ARI = round(mean(ARI),2)) %>%
  tidyr::pivot_wider(names_from = c(n_obs, Scenario), values_from = ARI) %>%
  # keep first 2 columns with model info, sort remaining cols with results by n
  dplyr::select(1,2, gtools::mixedorder(names(.)[3:length(names(.))])+2)
```

### KL divergence

```{r}
# KL
summary_table %>% 
  dplyr::filter(Tag == "sigmahyp") %>%
  dplyr::group_by(Model, Scenario, SM, n_obs) %>%
  dplyr::summarize(KL = round(mean(KL),2)) %>%
  tidyr::pivot_wider(names_from = c(n_obs, Scenario), values_from = KL) %>%
  # keep first 2 columns with model info, sort remaining cols with results by n
  dplyr::select(1,2, gtools::mixedorder(names(.)[3:length(names(.))])+2) 
```