---
title: "Conjugate Sampler - Unknown Variance (Fix b)"
author: "Jonathan Klus"
date: "2023-10-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load necessary functions
source("./Multivariate_DPMM_unknownvar_conjugate.R")
source("./posterior_helper_fxns.R")

# load R libraries
library(ggplot2)
library(plotly)
library(label.switching)
library(coda)
library(tidyr)
```

## EEE Sampler Setup

- $y_1, \dots, y_n \sim \sum_{k=1}^K\pi_kN(\mu_k, \sigma^2 I_J)$
- $\mu_1, \dots, \mu_K \sim N(\mu_0, r\sigma^2 I_J)$
- $\sigma^2 \sim IG(a,b)$
- $r \sim IG(g, h)$
- $b \sim Ga(d, f)$

## VEE Sampler Setup

- $y_1, \dots, y_n \sim \sum_{k=1}^K\pi_kN(\mu_k, \sigma^2_k I_J)$
- $\mu_1, \dots, \mu_K \sim N(\mu_0, r\sigma^2_k I_J)$
- $\sigma^2_1,\dots, \sigma^2_K \sim IG(a,b)$
- $r \sim IG(g, h)$
- $b \sim Ga(d, f)$

## EEE vs. VEE Sampler Examples

### 2D Example - well separated

Mixture of 3 multivariate normal densities all with diagonal variance structure
and $\sigma^2 = 10$. The true means are $(-20,20)$, $(20,-20)$, and $(0,0)$, and
$\sum_{j}n_j = 30$.

```{r, echo=FALSE}
################## simulate data ##############################################
w = c(0.4, 0.3, 0.3)
means = list(
  c(-20, 20),
  c(20, -20),
  c(0, 0)
)

var = diag(10, length(means[[1]])) # variances known, diagonal, and equal

assign = sample(x = 1:length(means), size = 30, replace = TRUE, prob = w)
y = lapply(X = assign,
           FUN = function(x){
             t(mvtnorm::rmvnorm(n = 1, mean = means[[x]], sigma = var))
           })
y_matrix = matrix(data = unlist(y), ncol = 2, byrow = TRUE)

data = data.frame(
  assign = assign,
  y1 = y_matrix[,1],
  y2 = y_matrix[,2]
)

ggplot(data = data, aes(x = y1, y = y2, label = rownames(data))) +
  #geom_point(color = assign) +
  #geom_text(size = 3, hjust = 0, nudge_x = 0.5, color = assign) +
  geom_text(size = 3, color = assign) +
  ggtitle("Simulated 2D Data and True Group Assignments") + 
  theme_classic()

# drop params used to generate data, except for var which is assumed known
rm(w, means)
```
#### EEE sampler 

```{r}
# sample
set.seed(516)
n_iter = 10^3
mu0 = matrix(round((colMeans(matrix(unlist(y), ncol = 2))),0), ncol = 1)
print(mu0)
ex1 = MVN_CRP_sampler_EEE(S = n_iter, y = y, alpha = 1, r = 10, 
                          a = 1, b = 50, d = 1, f = 10, 
                          g = 1, h = 50, 
                          sigma_hyperprior = FALSE, fix_r = FALSE, mu0 = mu0, 
                          k_init = 1, diag_weights = FALSE, 
                          verbose = TRUE, print_iter = 1000)
saveRDS(object = ex1, file = "./MCMC_Runs/conjEEEsamp_wellsep_a1_b50_d1_f20_g1_h50_rvaries.rds")

table(ex1$k)

# traceplot for number of components k.

make_k_traceplot(k = ex1$k, group_assign = ex1$group_assign)

ggplot(mapping = aes(x = 101:n_iter, y = unlist(ex1$vars)[101:n_iter])) +
  geom_line() +
  ggtitle("Traceplot of variance") +
  xlab("iter") +
  ylab("sigma^2") +
  theme_classic()

summary(unlist(ex1$vars)[101:n_iter])
summary(ex1$extra_params[101:n_iter,"b"])
summary(ex1$extra_params[101:n_iter,"r"])
```
#### VEE sampler 

```{r}
# sample
set.seed(516)
n_iter = 10^3
mu0 = matrix(round((colMeans(matrix(unlist(y), ncol = 2))),0), ncol = 1)
print(mu0)
ex1 = MVN_CRP_sampler_EEE(S = n_iter, y = y, alpha = 1, r = 10, 
                          a = 1, b = 20, g = 1, h = 50, 
                          sigma_hyperprior = FALSE, fix_r = FALSE, mu0 = mu0, 
                          k_init = 1, diag_weights = FALSE, 
                          verbose = TRUE, print_iter = 1000)
#saveRDS(object = ex1, file = "./MCMC_Runs/conjVEEsamp_wellsep_a1_b50_d1_f20_g1_h50_rvaries.rds")

table(ex1$k)

# traceplot for number of components k.

make_k_traceplot(k = ex1$k, group_assign = ex1$group_assign)

ggplot(mapping = aes(x = 101:n_iter, y = unlist(ex1$vars)[101:n_iter])) +
  geom_line() +
  ggtitle("Traceplot of variance") +
  xlab("iter") +
  ylab("sigma^2") +
  theme_classic()

summary(unlist(ex1$vars)[101:n_iter])
summary(ex1$extra_params[101:n_iter,"b"])
summary(ex1$extra_params[101:n_iter,"r"])
```

