---
title: "Conjugate Sampler - Unknown Variance"
author: "Jonathan Klus"
date: "2023-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load necessary functions
source("./Multivariate_DPMM_unknownvar_conjugate.R")
source("./posterior_helper_fxns.R")

# load R libraries
library(ggplot2)
library(plotly)
library(label.switching)
library(coda)
library(tidyr)
```

## EEE Sampler Setup

- $y_1, \dots, y_n \sim N(\mu, \sigma^2 I_J)$
- $\mu_1, \dots, \mu_K \sim N(\mu_0, r\sigma^2 I_J)$
- $\sigma^2 \sim IG(a,b)$
- $r \sim IG(g, h)$
- $b \sim Ga(d, f)$

### 2D Example - well separated

Mixture of 3 multivariate normal densities all with diagonal variance structure
and $\sigma^2 = 10$. The true means are $(-20,20)$, $(20,-20)$, and $(0,0)$, and
$\sum_{j}n_j = 30$.

```{r, echo=FALSE}
################## simulate data ##############################################
w = c(0.4, 0.3, 0.3)
means = list(
  c(-20, 20),
  c(20, -20),
  c(0, 0)
)

var = diag(10, length(means[[1]])) # variances known, diagonal, and equal

assign = sample(x = 1:length(means), size = 30, replace = TRUE, prob = w)
y = lapply(X = assign,
           FUN = function(x){
             t(mvtnorm::rmvnorm(n = 1, mean = means[[x]], sigma = var))
           })
y_matrix = matrix(data = unlist(y), ncol = 2, byrow = TRUE)

data = data.frame(
  assign = assign,
  y1 = y_matrix[,1],
  y2 = y_matrix[,2]
)

ggplot(data = data, aes(x = y1, y = y2, label = rownames(data))) +
  #geom_point(color = assign) +
  #geom_text(size = 3, hjust = 0, nudge_x = 0.5, color = assign) +
  geom_text(size = 3, color = assign) +
  ggtitle("Simulated 2D Data and True Group Assignments") + 
  theme_classic()

# drop params used to generate data, except for var which is assumed known
rm(w, means)
```

```{r}
# sample
set.seed(516)
n_iter = 10^4
mu0 = matrix(round((colMeans(matrix(unlist(y), ncol = 2))),0), ncol = 1)
print(mu0)
ex1 = MVN_CRP_sampler_EEE(S = n_iter, y = y, alpha = 1, r = 10, 
                          a = 1, b = 10, d = 1, f = 10, g = 2, h = 10, 
                          sigma_hyperprior = TRUE, fix_r = FALSE, mu0 = mu0, 
                          k_init = 1, diag_weights = FALSE, 
                          verbose = TRUE, print_iter = 100)
table(ex1$k)

# traceplot for number of components k.

make_k_traceplot(k = ex1$k, group_assign = ex1$group_assign)

sample(ex1$vars, size = 100)
  
# ggplot(mapping = aes(x = 1:n_iter, y = ex3$k)) + 
#   geom_line() +
#   ggtitle("Traceplot of no. components K") +
#   xlab("iter") + 
#   ylab("K") + 
#   theme_classic()

# # posterior inference before label switching fixed
# mean_list_by_k1 = list_params_by_k(draws = ex1$means, burn_in = 1000, param_type = "Mean", k_vec = ex1$k)
# var_list_by_k1 = list_params_by_k(draws = ex1$vars, burn_in = 1000, param_type = "Var", k_vec = ex1$k)
# #sigma22 is group 2, 2nd diagonal element
# prob_list_by_k1 = get_probs_by_k(probs = ex1$group_probs, 
#                                 n_groups = ex1$k, 
#                                 burn_in = 1000)
# group_assign_list_by_k1 = get_assign_by_k(assign = ex1$group_assign, 
#                                          n_groups = ex1$k, 
#                                          burn_in = 1000)
# ## deal with label switching for each k
# 
# ## in this case need to skip first result bc only one group was found -- can't 
# ## have label switching when there is only 1 group so the function fails
# 
# ## this ended up not being true here after some settings were changed --- need to program
# ## a more robust way to do this to avoid manually changing Stephens settings and potentially
# ## causing erros
# stephens_result1 = get_stephens_result(group_assign_list_by_k = group_assign_list_by_k1, 
#                                        prob_list_by_k = prob_list_by_k1)
# 
# ## now reorder means to deal with label switching and redo posterior inference and 
# ## traceplots
# mean_list_by_k_stephens1 = list_params_by_k(draws = ex1$means, k_vec = ex1$k,
#                                            burn_in = 1000,
#                                            relabel = TRUE,
#                                            permutation = stephens_result1, 
#                                            param_type = "Mean")
# var_list_by_k_stephens1 = list_params_by_k(draws = ex1$vars, k_vec = ex1$k,
#                                            burn_in = 1000,
#                                            relabel = TRUE,
#                                            permutation = stephens_result1,
#                                            param_type = "Var")
# 
# # posterior means
# make_postsum(mcmc_df = mean_list_by_k_stephens1[[1]], digits = 2)
# make_postsum(mcmc_df = mean_list_by_k_stephens1[[2]], digits = 2)
# make_postsum(mcmc_df = mean_list_by_k_stephens1[[3]], digits = 2)
# 
# # posterior vars
# make_postsum(mcmc_df = var_list_by_k_stephens1[[1]], digits = 2)
# make_postsum(mcmc_df = var_list_by_k_stephens1[[2]], digits = 2)
# make_postsum(mcmc_df = var_list_by_k_stephens1[[3]], digits = 2)
# 
# 
# 
# 
# # empirical means & vars
# # emp_mean_list_by_k_stephens3 = list_params_by_k(draws = ex3$emp_means, k_vec = ex3$k,
# #                                            burn_in = 1000,
# #                                            relabel = TRUE,
# #                                            permutation = stephens_result3, 
# #                                            param_type = "Mean")
# emp_var_list_by_k_stephens1 = list_params_by_k(draws = ex1$emp_vars, k_vec = ex1$k,
#                                            burn_in = 1000,
#                                            relabel = TRUE,
#                                            permutation = stephens_result1,
#                                            param_type = "Var")
# 
# # empirical vars
# make_postsum(mcmc_df = emp_var_list_by_k_stephens1[[1]], digits = 2)
# make_postsum(mcmc_df = emp_var_list_by_k_stephens1[[2]], digits = 2)
# make_postsum(mcmc_df = emp_var_list_by_k_stephens1[[3]], digits = 2)
# 
# make_traceplot(param_list_by_k = mean_list_by_k_stephens1, 
#                k = 3, p = 2,
#                component_no = 1,
#                param_type = "Mean")
# make_traceplot(param_list_by_k = mean_list_by_k_stephens1, 
#                k = 3, p = 2,
#                component_no = 2,
#                param_type = "Mean")
# make_traceplot(param_list_by_k = var_list_by_k_stephens1, 
#                     k = 3, 
#                     component_no = 1,
#                     p=2,
#                     param_type = "Var")
# make_traceplot(param_list_by_k = var_list_by_k_stephens1, 
#                     k = 3, 
#                     component_no = 2,
#                     p=2,
#                     param_type = "Var")
# 
# make_traceplot(param_list_by_k = mean_list_by_k_stephens1, 
#                k = 4, p = 2,
#                component_no = 1,
#                param_type = "Mean")
# make_traceplot(param_list_by_k = mean_list_by_k_stephens1, 
#                k = 4, p = 2,
#                component_no = 2,
#                param_type = "Mean")
# 
# make_traceplot(param_list_by_k = var_list_by_k_stephens1, 
#                     k = 4, 
#                     component_no = 1,
#                     p=2,
#                     param_type = "Var")
# make_traceplot(param_list_by_k = var_list_by_k_stephens1, 
#                     k = 4, 
#                     component_no = 2,
#                     p=2,
#                     param_type = "Var")
# 
# make_traceplot(param_list_by_k = mean_list_by_k_stephens1, 
#                k = 5, p = 2,
#                component_no = 1,
#                param_type = "Mean")
# make_traceplot(param_list_by_k = mean_list_by_k_stephens1, 
#                k = 5, p = 2,
#                component_no = 2,
#                param_type = "Mean")
# make_traceplot(param_list_by_k = var_list_by_k_stephens1, 
#                     k = 5, 
#                     component_no = 1,
#                     p=2,
#                     param_type = "Var")
# make_traceplot(param_list_by_k = var_list_by_k_stephens1, 
#                     k = 5, 
#                     component_no = 2,
#                     p=2,
#                     param_type = "Var")

```

## 2D Example - close together

Mixture of 3 multivariate normal densities all with diagonal variance structure
and $\sigma^2 = 10$. The true means are $(10,10)$, $(0,-5)$, , $(12,-2)$, and $(0,10)$, and
$\sum_{j}n_j = 50$.

```{r, echo=FALSE}
################## simulate data ##############################################
set.seed(1234)
w = c(.35, .25, .4)
means = list(
  #c(10, 10),
  c(0, -5),
  c(0, 10),
  c(12,2)
)

var = diag(9, length(means[[1]])) # variances known, diagonal, and equal

assign = sample(x = 1:length(means), size = 50, replace = TRUE, prob = w)
y = lapply(X = assign,
           FUN = function(x){
             t(mvtnorm::rmvnorm(n = 1, mean = means[[x]], sigma = var))
           })
y_matrix = matrix(data = unlist(y), ncol = 2, byrow = TRUE)

data = data.frame(
  assign = assign,
  y1 = y_matrix[,1],
  y2 = y_matrix[,2]
)

# empirical means and variances
data %>%
  group_by(assign) %>%
  dplyr::summarize(mean(y1), mean(y2), var(y1), var(y2))

ggplot(data = data, aes(x = y1, y = y2)) +
  geom_point(color = assign) +
  ggtitle("Simulated 2D Data and True Group Assignments") + 
  theme_classic()

# drop params used to generate data, except for var which is assumed known
rm(w, means)
```


```{r,  message=FALSE}
mu0 = matrix(round((colMeans(matrix(unlist(y), ncol = 2))),0), ncol = 1)
print(mu0)

# sample
set.seed(516)
n_iter = 10^4
ex1 = MVN_CRP_sampler_EEE(S = n_iter, y = y, alpha = 1, r = 10, 
                          a = 1, b = 10, d = 1, f = 10, g = 2, h = 10, 
                          sigma_hyperprior = TRUE, fix_r = FALSE, mu0 = mu0, 
                          k_init = 1, diag_weights = FALSE, 
                          verbose = TRUE, print_iter = 100)
table(ex1$k)

# traceplot for number of components k.

make_k_traceplot(k = ex1$k, group_assign = ex1$group_assign)

sample(ex1$vars, size = 100)
```

